---
title: "R Notebook"
output: html_notebook
---

# MR WorkFlow:

```{r}
library(TwoSampleMR) 
library(tidyverse)  

```

### 1. Importing Summary Statistics

```{r}
#Load location of files
SB <- system.file("extdata", "SBP_Sig.txt", package = "TwoSampleMR")
HF <- system.file("extdata", "HFdata.txt", package = "TwoSampleMR")
HF
```

```{r}
#use read_exposure function to read exposure data 
SB_exp_data <- read_exposure_data(SB)
SB_exp_data
```

### 2. Performing LD Clumping of Exposure Data, and Harmonisation

```{r}

#CLUMPING EXPOSURE DATA - ensure independent SNPs selected
options(ieugwasr_api = 'gwas-api.mrcieu.ac.uk/')
SB_clump <- clump_data(SB_exp_data, clump_r2 = 0.01, pop = "EUR")



#Read outcome data for the clumped exposure SNPs
HF_out_data <- read_outcome_data(HF, snps = SB_clump$SNP)



#PERFORMING HARMONISATION
hd <- harmonise_data(SB_clump, HF_out_data, action = 2) %>% as_tibble() %>%
  filter(pval.outcome > 5e-8, eaf.outcome > 0.05, eaf.exposure > 0.05)
#action 3 = Correct the strand for non-palindromic SNPs, but drop all palindromic SNPs


```

### 3. Running MR Analysis

```{r}
results <- mr(hd)
results$Lower_CI <- results$b - 1.96 * results$se
results$Upper_CI <- results$b + 1.96 * results$se
results

#ALL RESULTS SHOW SIGNIFICANT EVIDENCE TO ACCEPT NULL HYPOTHESIS OF THERE BEING NO CAUSAL EFFECT. 
```

# Sensitivity Analyses:

## 1. Pleiotropy Test:

```{r}
mr_pleiotropy_test(hd)
#p_value much larger than 0.05 displays no strong evidence of directional horizontal pleiotropy in this analysis
```

## 2. Heterogeneity Test:

```{r}
heterogeneity_test <- mr_heterogeneity(hd)
heterogeneity_test


```

# Visual Plots:

## 1. Scatter Plot:

```{r}
plot1 <- mr_scatter_plot(results,hd)
plot1
```

## 2. Forest plot:

```{r}
res_single <- mr_singlesnp(hd)
plot2 <- mr_forest_plot(res_single)
plot2
```

## 3. Leave-One-Out Plot:

```{r}
res_loo <- mr_leaveoneout(hd)
plot3 <- mr_leaveoneout_plot(res_loo)
plot3
```

## 4. Funnel Plot: (Visually Inspect Horizontal Pleiotropy)

```{r}
plot4 <- mr_funnel_plot(res_single)
plot4
#Asymmetry can indicate invalid instruments
```

# MendelianRandomization Package:

### 1. IVW estimate using mr_ivw function:

```{r}
mrdata <- mr_input(bx = hd$beta.exposure, bxse = hd$se.exposure, by = hd$beta.outcome, byse = hd$se.outcome, snps = hd$SNP)

mrresult <- result_ivw <- MendelianRandomization::mr_ivw(mrdata, robust = TRUE, penalized = TRUE)
mrresult
```
